Estimation of Spatiotemporal Integro-Difference Equations for Dynamical Neural Field Models

### 2009-11-18 ltx/BrainIDE.tex
added yet another SS derivation. I made an error with previous derivation where I separated a spatial and time variable incorrectly

\author{Dean~Freestone,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Parham~Aram,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Michael~Dewar,~\IEEEmembership{Member,~IEEE,}
        Kenneth~Scerri,~\IEEEmembership{Member,~IEEE,}
        and Visakan~Kadirkamanathan,~\IEEEmembership{Member,~IEEE,}
        and~other~Bosses/Supervisors,~\IEEEmembership{Member,~IEEE.}% <-this % stops a space% <-this % stops a space

@@

%\subsection{IDE Formulation}
%For the state-space derivation, to create an estimator, the model can be reformulated like so
%\begin{equation}\label{new_spatiotemporal_kernels}
%\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {k\left( {r - r'} \right)\tilde v\left( {r',t} \right) - \alpha \delta \left( {r - r'} \right)\bar v\left( {r,t} \right)dr'}  + e\left( {r,t} \right)
%\begin{equation}\label{define_v_tilde}
%\tilde v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
%   {v\left( {r,t} \right)} & {v\left( {r,t} \right) \ge {v_0}}  \\
%\begin{equation}\label{define_v_bar}
%\bar v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
%   {v\left( {r,t} \right)} & {v\left( {r,t} \right) < {v_0}}  \\
%\includegraphics[scale=0.4]{ExampleFieldDecomposition.eps}
%\caption[ExampleFieldDecomposition]{Illustration of a one dimensional field decomposition using 5 Gaussian basis functions, each weighted an element of the state vector. Each basis function has a width of 0.01 and is uniformly positioned with a separation of 0.25 (with arbitrary units).}
%\label{ExampleFieldDecomposition}

@@

 v_{t+1}(\mathbf s)=\int_{\mathcal S} \alpha \delta(\mathbf s-\mathbf r)\bar v_t(\mathbf r)d \mathbf r+\int_{\mathcal S} w(\mathbf s -\mathbf r)\tilde v_t(\mathbf r)d\mathbf r+e_t(\mathbf s)

@@

\tilde v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_t=\nu_0 + \boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_t

@@

 \boldsymbol \phi^{\top}(\mathbf s)\bar{\mathbf x}_{t+1}+\boldsymbol \phi^{\top}(\mathbf s)\tilde{\mathbf x}_{t+1}=
\int_\mathcal{S}\alpha\delta(\mathbf s- \mathbf r)\boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_td\mathbf r+\int_\mathcal{S}\boldsymbol \theta^{\top}\boldsymbol\psi(\mathbf s -\mathbf r)\boldsymbol\phi^{\top}(\mathbf r)\tilde{\mathbf x}_td\mathbf r+e_t(\mathbf s)

@@

 \boldsymbol \phi^{\top}(\mathbf s)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf s)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\int_{\mathcal S}\boldsymbol\psi(\mathbf s -\mathbf r)\boldsymbol\phi^{\top}(\mathbf r)\tilde{\mathbf x}_td\mathbf r+e_t(\mathbf s)

@@

 \boldsymbol\Phi(\mathbf s)=\int_{\mathcal S}\boldsymbol\psi(\mathbf s -\mathbf r)\boldsymbol\phi^{\top}(\mathbf r)d\mathbf r

@@

 \boldsymbol \phi^{\top}(\mathbf s)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf s)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf s)\tilde{\mathbf x}_t+e_t(\mathbf s)

@@

Premultiplying (\ref{eq:cortexfieldwithPhi}) by $\boldsymbol\phi(\mathbf s) $, and integrating over $\mathbf s$

@@

\int_{\mathcal S} \boldsymbol\phi(\mathbf s)\boldsymbol \phi^{\top}(\mathbf s)d\mathbf s(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\int_{\mathcal S} \boldsymbol\phi(\mathbf s)\boldsymbol \phi^{\top}(\mathbf s)d\mathbf s \bar{\mathbf x}_{t}+\int_{\mathcal S} \boldsymbol\phi(\mathbf s)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf s)d\mathbf s\tilde{\mathbf x}_t+\int_{\mathcal S} \boldsymbol\phi(\mathbf s)e_t(\mathbf s)d\mathbf s

@@

 \boldsymbol\Psi_x(\mathbf s)=\int_{\mathcal S}\boldsymbol\phi(\mathbf s)\boldsymbol\phi^{\top}(\mathbf s)d\mathbf s

@@

  \boldsymbol\Psi_{\theta}(\mathbf s)=\int_{\mathcal S} \boldsymbol\phi(\mathbf s)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf s)d\mathbf s

@@

\boldsymbol\Psi_x(\mathbf s)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\boldsymbol\Psi_x(\mathbf s) \bar{\mathbf x}_{t}+\boldsymbol\Psi_{\theta}(\mathbf s)\tilde{\mathbf x}_t+\int_{\mathcal S} \boldsymbol\phi(\mathbf s)e_t(\mathbf s)d\mathbf s

@@

\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1}=\alpha\bar{\mathbf x}_{t}+\boldsymbol\Psi_x^{-1}\boldsymbol\Psi_{\theta}\tilde{\mathbf x}_t+\boldsymbol\Psi_x^{-1}\int_{\mathcal S} \boldsymbol\phi(\mathbf s)e_t(\mathbf s)d\mathbf s

@@

 \mathbf y_t(\mathbf s)&=&\\
&=&\int_{\mathcal S}H(\mathbf s-\mathbf r)v_t(\mathbf r)d\mathbf r+\boldsymbol\epsilon_t \\
&=&\int_{\mathcal S}H(\mathbf s-\mathbf r)\bar v_t(\mathbf r)d\mathbf r+\int_{\mathcal S}H(\mathbf s-\mathbf r)\tilde v_t(\mathbf r)d\mathbf r+\epsilon_t

@@

 we decompose $ H(\mathbf s-\mathbf r)$ as
 \begin{equation}
  H(\mathbf s-\mathbf r)= \mathbf c^{\top}\mu(\mathbf s-\mathbf r)
\label{eq:decomposedobservationkernel}
\end{equation}
and substituting  (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:decomposedobservationkernel}) in (\ref{eq:observationequation}) we get
\begin{equation}
 \mathbf y_t(\mathbf s)=\int_{\mathcal S}\mathbf c^{\top}\mu(\mathbf s-\mathbf r)\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}d\mathbf r+\int_{\mathcal S}\mathbf c^{\top}\mu(\mathbf s-\mathbf r)\phi^{\top}(\mathbf r)\tilde{\mathbf x}_{t}d\mathbf r+\epsilon_t
\end{equation}
and finally we have
\begin{equation}
\mathbf y_t(\mathbf s)= \begin{bmatrix}\mathbf c^{\top}\mathbf M(\mathbf s)&\mathbf c^{\top}\mathbf M(\mathbf s)\end{bmatrix}\underline{\mathbf x}_t+\epsilon_t
\end{equation}
where
\begin{equation}
 \mathbf M(\mathbf s)=\int_{\mathcal S}\mu(\mathbf s-\mathbf r)\phi^{\top}(\mathbf r)d\mathbf r
\end{equation}
Next we shall demonstrate the stability of the model and in the following section the estimator for the state vector and connectivity structure is derived.

\subsection{State-Space Neural Field Model}
In this section we shall reconstruct the model in a state-space form that is a novel representation of the neural field equations. The benefit of restating the model in this format is that it becomes easier to simulate and allows the connectivity to be inferred from data. We begin with the model in the form
\begin{eqnarray}\label{InitialModel}
% \nonumber to remove numbering (before each equation)
  {v_{t + 1}}\left( r \right) &=& \int_\Omega  {k\left( {r - r'} \right)p\left( {{v_t}\left( {r'} \right)} \right) - \alpha v\left( {r,t} \right)dr'}  + \varepsilon \left( {r,t} \right) \nonumber \\
   &=& \int_\Omega  {k\left( {r - r'} \right)p\left( {{v_t}\left( {r'} \right)} \right)dr'}  - \alpha v\left( {r,t} \right) + \varepsilon \left( {r,t} \right).
Next we defining the basis functions the decompositions of the field and connectivity kernel, where $\phi(s)$ is a vector is Gaussian basis functions, $x_t$ is the state, $\psi(r-r')$ is a vector of three connectivity basis functions and $\theta$ is the corresponding weight of the connectivity
\begin{equation}\label{FieldDecomp}
   {v_t}\left( r \right) = {\phi ^T}\left( r \right){x_t}
\begin{equation}\label{KernelDecomp}
k\left( {r - r'} \right) = {\theta ^T}\psi \left( {r - r'} \right).
Assume we know $\phi(r')$ and it is constant, and the form of the activation function $p(.)$ is known, and $\phi(r')$ is independent of $x_t$, then we can uniquely define the function $\rho(.)$ such that
\begin{eqnarray}\label{NonlinearState}
% \nonumber to remove numbering (before each equation)
  p\left( {{v_t}\left( {r'} \right)} \right)  &=& p\left( {{\phi ^T}\left( {r'} \right){x_t}} \right) \nonumber \\
   &=& \phi ^T\left(r'\right)\rho \left( {{x_t}} \right).
Making substitutions of~\ref{FieldDecomp}, \ref{KernelDecomp} and \ref{NonlinearState} into \ref{InitialModel} we get
\begin{equation}\label{ModelDecomposed}
{\phi ^T}\left( r \right){x_{t + 1}} = {\theta ^T}\int_\Omega  {\psi \left( {r - r'} \right){\phi ^T}\left( {r'}\right)dr'} \rho \left( {{x_t}} \right) - \alpha {\phi ^T}\left( r \right){x_t} + \varepsilon \left( {r,t} \right)
\begin{equation}\label{DefPhi}
\Phi \left( r \right) = \int_\Omega  {\psi \left( {r - r'} \right){\phi ^T}\left( {r'} \right)dr'}.
Substituting~\ref{DefPhi} into \ref{ModelDecomposed} we get
\begin{equation}\label{Sub2}
{\phi ^T}\left( r \right){x_{t + 1}} = {\theta ^T}\Phi \left( r \right)\rho \left( {{x_t}} \right) - \alpha {\phi ^T}\left( r \right){x_t} + \varepsilon \left( {r,t} \right).
Cross multiplying by $\phi(r)$ in integrating over $r$ gives
\begin{equation}\label{CorssMultAndInt}
\int_\Omega  {\phi \left( r \right){\phi ^T}\left( r \right)dr} {x_{t + 1}} = {\theta ^T}\int_\Omega  {\phi \left( r \right)\Phi \left( r \right)dr} \rho \left( {{x_t}} \right) - \alpha \int_\Omega  {\phi \left( r \right){\phi ^T}\left( r \right)dr} {x_t} + \int_\Omega  {\phi \left( r \right)\varepsilon \left( {r,t} \right)dr}.
\begin{equation}\label{DefPsix}
{\Psi _x} = \int_\Omega  {\phi \left( r \right){\phi ^T}\left( r \right)dr}
\begin{equation}\label{DefPsitheta}
{\Psi _\theta } = \int_\Omega  {\phi \left( r \right)\Phi \left( r \right)dr}.
Substituting~\ref{DefPsix} and \ref{DefPsitheta} into \ref{CorssMultAndInt} we get 
\begin{equation}\label{Subs3}
{\Psi _x}{x_{t + 1}} = {\theta ^T}{\Psi _\theta }\rho \left( {{x_t}} \right) - \alpha {\Psi _x}{x_t} + \int_\Omega  {\phi \left( r \right)\varepsilon \left( {r,t} \right)dr}.
Now we isolate the state by inverting $\Psi_x$
\begin{equation}\label{IsolateState}
{x_{t + 1}} = {\Psi _x}^{ - 1}{\theta ^T}{\Psi _\theta }\rho \left( {{x_t}} \right) - \alpha {x_t} + {\Psi _x}^{ - 1}\int_\Omega  {\phi \left( r \right)\varepsilon \left( {r,t} \right)dr}.
This gives the state-space model
\begin{equation}\label{StateSpaceModel}
{x_{t + 1}} = f\left( {x,\theta } \right) + {\Psi _x}^{ - 1}\int_\Omega  {\phi \left( r \right)\varepsilon \left( {r,t} \right)dr}.

@@

\begin{IEEEbiography}{Dean Freestone}
Biography text here.
\end{IEEEbiography}

>>>>

\author{Dean~R.~Freestone,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Parham~Aram,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Michael~Dewar,~\IEEEmembership{Member,~IEEE,}
        Kenneth~Scerri,~\IEEEmembership{Member,~IEEE,}
        Visakan~Kadirkamanathan,~\IEEEmembership{Member,~IEEE,}
        David~B.~Grayden,~\IEEEmembership{Member,~IEEE,}
        Anthony~N.~Burkitt,~\IEEEmembership{Member,~IEEE,}
        and~other~Bosses/Supervisors,~\IEEEmembership{Member,~IEEE.}% <-this % stops a space% <-this % stops a space

@@

\subsection{IDE Formulation}
For the state-space derivation, to create an estimator, the model can be reformulated like so
\begin{equation}\label{new_spatiotemporal_kernels}
\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {k\left( {r - r'} \right)\tilde v\left( {r',t} \right) - \alpha \delta \left( {r - r'} \right)\bar v\left( {r,t} \right)dr'}  + e\left( {r,t} \right)
\begin{equation}\label{define_v_tilde}
\tilde v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
   {v\left( {r,t} \right)} & {v\left( {r,t} \right) \ge {v_0}}  \\
   0 & {otherwise}  \\
\end{array}} \right.
\begin{equation}\label{define_v_bar}
\bar v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
   {v\left( {r,t} \right)} & {v\left( {r,t} \right) < {v_0}}  \\
   0 & {otherwise}  \\
\end{array}} \right.
\begin{figure*}
\centering
\includegraphics[scale=0.3]{ExampleFieldDecomposition.eps}
\caption[ExampleFieldDecomposition]{Illustration of a one dimensional field decomposition using 5 Gaussian basis functions, each weighted an element of the state vector. Each basis function has a width of 0.01 and is uniformly positioned with a separation of 0.25 (with arbitrary units).}
\label{ExampleFieldDecomposition}
\end{figure*}

@@

 v_{t+1}(\mathbf r)=\int_{\Omega} \alpha \delta(\mathbf r-\mathbf r')\bar v_t(\mathbf r)d \mathbf r'+\int_{\Omega} w(\mathbf r -\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+e_t(\mathbf r)

@@

\tilde v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_t=\nu_0

@@

 \boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_{t+1}+\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_{t+1}=
\int_\Omega\alpha\delta(\mathbf r- \mathbf r')\boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_td\mathbf r'+\int_\Omega\boldsymbol \theta^{\top}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)

@@

 \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)

@@

 \boldsymbol\Phi(\mathbf r)=\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')d\mathbf r'

@@

 \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)\tilde{\mathbf x}_t+e_t(\mathbf r)

@@

Premultiplying (\ref{eq:cortexfieldwithPhi}) by $\boldsymbol\phi(\mathbf r) $, and integrating over $\mathbf r$

@@

\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf s(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf r \bar{\mathbf x}_{t}+\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r

@@

 \boldsymbol\Psi_x(\mathbf r)=\int_{\Omega}\boldsymbol\phi(\mathbf r)\boldsymbol\phi^{\top}(\mathbf r)d\mathbf r

@@

  \boldsymbol\Psi_{\theta}(\mathbf r)=\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r

@@

\boldsymbol\Psi_x(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\boldsymbol\Psi_x(\mathbf r) \bar{\mathbf x}_{t}+\boldsymbol\Psi_{\theta}(\mathbf r)\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r

@@

\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1}=\alpha\bar{\mathbf x}_{t}+\boldsymbol\Psi_x^{-1}\boldsymbol\Psi_{\theta}\tilde{\mathbf x}_t+\boldsymbol\Psi_x^{-1}\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r

@@

 \mathbf y_t(\mathbf r)&=&\\
&=&\int_{\Omega}H(\mathbf r-\mathbf r')v_t(\mathbf r')d\mathbf r'+\boldsymbol\epsilon_t \\
&=&\int_{\Omega}H(\mathbf r-\mathbf r')\bar v_t(\mathbf r')d\mathbf r'+\int_{\Omega}H(\mathbf r-\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+\epsilon_t

@@

 we decompose $ H(\mathbf r-\mathbf r')$ as
 \begin{equation}
  H(\mathbf r-\mathbf r')= \mathbf c^{\top}\mu(\mathbf r-\mathbf r')
\label{eq:decomposedobservationkernel}
\end{equation}
and substituting  (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:decomposedobservationkernel}) in (\ref{eq:observationequation}) we get
\begin{equation}
 \mathbf y_t(\mathbf r)=\int_{\Omega}\mathbf c^{\top}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')\bar{\mathbf x}_{t}d\mathbf r'+\int_{\Omega}\mathbf c^{\top}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')\tilde{\mathbf x}_{t}d\mathbf r'+\epsilon_t
\end{equation}
and finally we have
\begin{equation}
\mathbf y_t(\mathbf r)= \begin{bmatrix}\mathbf c^{\top}\mathbf M(\mathbf r)&\mathbf c^{\top}\mathbf M(\mathbf r)\end{bmatrix}\underline{\mathbf x}_t+\epsilon_t
\end{equation}
where
\begin{equation}
 \mathbf M(\mathbf r)=\int_{\Omega}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')d\mathbf r'
\end{equation}
Next we shall demonstrate the stability of the model and in the following section the estimator for the state vector and connectivity structure is derived.

%\subsection{State-Space Neural Field Model}
%In this section we shall reconstruct the model in a state-space form that is a novel representation of the neural field equations. The benefit of restating the model in this format is that it becomes easier to simulate and allows the connectivity to be inferred from data. We begin with the model in the form
%\begin{eqnarray}\label{InitialModel}
%  {v_{t + 1}}\left( r \right) &=& \int_\Omega  {k\left( {r - r'} \right)p\left( {{v_t}\left( {r'} \right)} \right) - \alpha v\left( {r,t} \right)dr'}  + \varepsilon_t \left( {r} \right) \nonumber \\
%   &=& \int_\Omega  {k\left( {r - r'} \right)p\left( {{v_t}\left( {r'} \right)} \right)dr'}  - \alpha v\left( {r,t} \right) + \varepsilon_t \left( {r} \right).
%Next we defining the basis functions the decompositions of the field and connectivity kernel, where $\phi(s)$ is a vector is Gaussian basis functions, $x_t$ is the state, $\psi(r-r')$ is a vector of three connectivity basis functions and $\theta$ is the corresponding weight of the connectivity
%\begin{equation}\label{FieldDecomp}
%   {v_t}\left( r \right) = {\phi ^T}\left( r \right){x_t}
%\begin{equation}\label{KernelDecomp}
%k\left( {r - r'} \right) = {\theta ^T}\psi \left( {r - r'} \right).
%Assume we know $\phi(r')$ and it is constant, and the form of the activation function $p(.)$ is known, and $\phi(r')$ is independent of $x_t$, then we can uniquely define the function $\rho(.)$ such that
%\begin{eqnarray}\label{NonlinearState}
%  p\left( {{v_t}\left( {r'} \right)} \right)  &=& p\left( {{\phi ^T}\left( {r'} \right){x_t}} \right) \nonumber \\
%   &=& \phi ^T\left(r'\right)\rho \left( {{x_t}} \right).
%Making substitutions of~\ref{FieldDecomp}, \ref{KernelDecomp} and \ref{NonlinearState} into \ref{InitialModel} we get
%\begin{equation}\label{ModelDecomposed}
%{\phi ^T}\left( r \right){x_{t + 1}} = {\theta ^T}\int_\Omega  {\psi \left( {r - r'} \right){\phi ^T}\left( {r'}\right)dr'} \rho \left( {{x_t}} \right) - \alpha {\phi ^T}\left( r \right){x_t} + \varepsilon_t \left( {r} \right)
%Now we define
%\begin{equation}\label{DefPhi}
%\Phi \left( r \right) = \int_\Omega  {\psi \left( {r - r'} \right){\phi ^T}\left( {r'} \right)dr'}.
%Substituting~\ref{DefPhi} into \ref{ModelDecomposed} we get
%\begin{equation}\label{Sub2}
%{\phi ^T}\left( r \right){x_{t + 1}} = {\theta ^T}\Phi \left( r \right)\rho \left( {{x_t}} \right) - \alpha {\phi ^T}\left( r \right){x_t} + \varepsilon_t \left( {r} \right).
%Cross multiplying by $\phi(r)$ in integrating over $r$ gives
%\begin{equation}\label{CorssMultAndInt}
%\int_\Omega  {\phi \left( r \right){\phi ^T}\left( r \right)dr} {x_{t + 1}} = {\theta ^T}\int_\Omega  {\phi \left( r \right)\Phi \left( r \right)dr} \rho \left( {{x_t}} \right) - \alpha \int_\Omega  {\phi \left( r \right){\phi ^T}\left( r \right)dr} {x_t} + \int_\Omega  {\phi \left( r \right)\varepsilon_t \left( {r} \right)dr}.
%Now we define
%\begin{equation}\label{DefPsix}
%{\Psi _x} = \int_\Omega  {\phi \left( r \right){\phi ^T}\left( r \right)dr}
%\begin{equation}\label{DefPsitheta}
%{\Psi _\theta } = \int_\Omega  {\phi \left( r \right)\Phi \left( r \right)dr}.
%Substituting~\ref{DefPsix} and \ref{DefPsitheta} into \ref{CorssMultAndInt} we get
%\begin{equation}\label{Subs3}
%{\Psi _x}{x_{t + 1}} = {\theta ^T}{\Psi _\theta }\rho \left( {{x_t}} \right) - \alpha {\Psi _x}{x_t} + \int_\Omega  {\phi \left( r \right)\varepsilon_t \left( {r} \right)dr}.
%Now we isolate the state by inverting $\Psi_x$
%\begin{equation}\label{IsolateState}
%{x_{t + 1}} = {\Psi _x}^{ - 1}{\theta ^T}{\Psi _\theta }\rho \left( {{x_t}} \right) - \alpha {x_t} + {\Psi _x}^{ - 1}\int_\Omega  {\phi \left( r \right)\varepsilon_t \left( {r} \right)dr}.
%This gives the state-space model
%\begin{equation}\label{StateSpaceModel}
%{x_{t + 1}} = f\left( {x,\theta } \right) + {\Psi _x}^{ - 1}\int_\Omega  {\phi \left( r \right)\varepsilon_t \left( {r} \right)dr}.

\section{State-Space Model}
If we are going to use a particle filter or a extended Kalman filter we might as well go for this case. If the temporal and spatial part of the nonlinearity can not be separated, than this is the best choice.
\begin{equation}\label{SystemEq}
{v_{t + 1}}\left( {\bf{r}} \right) = \int_\Omega  {w\left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}}  - \alpha {v_t}\left( {\bf{r}} \right) + {\varepsilon _t}\left( {\bf{r}} \right)
\end{equation}
Defining
\begin{equation}\label{CombiningKernels}
k\left( {{\bf{r}} - {\bf{r'}}} \right) = w\left( {{\bf{r}} - {\bf{r'}}} \right) - \hat{\alpha}\delta \left( {{\bf{r}} - {\bf{r'}}} \right).
\end{equation}
Making substitution we get
\begin{equation}\label{OneKernelModel}
{v_{t + 1}}\left( {\bf{r}} \right) = \int_\Omega  {k\left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}}  + {\varepsilon _t}\left( {\bf{r}} \right).
\end{equation}
Defining
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}
\end{equation}
\begin{equation}\label{DefKernelDecomp}
k\left( {{\bf{r}} - {\bf{r'}}} \right) = {\theta ^T}\psi \left( {{\bf{r}} - {\bf{r'}}} \right)
\end{equation}
Now we have four basis functions (the fourth being the delta/local connectivity) and four parameters to estimate. Making substitutions
\begin{equation}\label{DecompModel}
{\phi ^T}\left( {\bf{r}} \right){x_t} = {\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}}  + {\varepsilon _t}\left( {\bf{r}} \right)
\end{equation}
Cross multiply and integrate over $\bf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_t}}  = \int_\Omega  {\phi \left( {\bf{r}} \right){\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  + \int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
Defining
\begin{equation}\label{DefPsi}
\Psi  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
Making substitution
\begin{equation}\label{SS Model}
{x_t} = {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  + {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
\begin{equation}\label{AbbrevSSModel}
{x_{t + 1}} = f\left( {x;\theta } \right) + {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}

@@

\begin{IEEEbiography}{Dean R. Freestone}
Biography text here.
\end{IEEEbiography}
